Parallel Acceleration of Fixed Complexity Sphere Decoder for Large MIMO Uplink System Based on CUDA

### 2015-08-06 Report April 2014/draft_tradeoff_original.tex
Reorganized Report Record

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[12pt,a4paper,final]{article}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
\usepackage[latin1]{inputenc}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
\usepackage{psfrag}
\usepackage{color}
\usepackage{epsfig}
\usepackage{appendix}
\usepackage{caption}
\usepackage{cite}
\usepackage{ifpdf}
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{stfloats}
\usepackage{url}
\usepackage{fixltx2e}
\usepackage{setspace} 
\usepackage{diagbox}

@@

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Report}

@@

\section{System Model}\label{system}
% no \IEEEPARstart
%This demo file is intended to serve as a ``starter file''
%for IEEE conference papers produced under \LaTeX\ using
%IEEEtran.cls version 1.8 and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
%I wish you the best of success.
%\hfill mds
%\hfill December 27, 2012
Consider a MIMO system with $N_r$ receive antennas and $N_t$ transmit antennas, where $N_r$ is large, for simplification, we assume $N_r\to\infty$, the corresponding discrete time model is given:
\begin{equation}
y=\mathbf{H}s+n
\end{equation}
$s\in \mathbb{C}^{N_{t}\times 1}$ is the transmit symbol vector, $y\in \mathbb{C}^{N_{r}\times 1}$ is the receive symbol vector, $\mathbf{H}\in \mathbb{C}^{N_{r}\times N_{t}}$ is the channel propagation matrix with independent identity distributed(i.i.d) Circularly Symmetric Complex Gaussian(CSCG) zero mean elements of unit variance (Rayleigh fading), $n\in \mathbb{C}^{N_{r}\times 1}$ is the additive white Gaussian noise (AWGN) with zero mean elements of variance $\sigma^{2}$, $\varepsilon(ss^{H})=E_{s}\mathbf{I}_{Nt}$, $E_s$ denotes the average symbol energy, $\varepsilon$ denotes the expectation operation and $M$ denotes the constellation size.

@@

y_{d}=\mathbf{H}_{j}s_{d}+n

@@

\hat{s_{d}}=\mathbf{G}_{j}y_{d}=\mathbf{G}_{j}\mathbf{H}_{j}s_{d}+\mathbf{G}_{j}n\label{equation0}

@@

[\hat{s_{d}}]_{m}=(gh)_{m}s_{d}+g_{m}n=g_{m}h_{m}[s_{d}]_{m}+\sum_{f\neq m}g_{m}h_{f}[s_{d}]_{f}+g_{m}n\label{equation01}

@@

where $[\hat{s_{d}}]_{m}$ denotes the $m$th element of $\hat{s_{d}}$, $(gh)_{m}$ denotes the $m$th row of $\mathbf{G}\mathbf{H}_{j}$, $g_{m}$ denotes the $m$th row of $\mathbf{G}_{j}$ and $h_{m}$ denotes the $m$th column of $\mathbf{H}_{j}$, The post processing SINR of $m$th sub datastream and $j$th selected antenna subset is defined as\cite{3}:

@@

\varepsilon[|g_{m}n|^{2}]= \varepsilon[(g_{m}n)(g_{m}n)^{H}]=\varepsilon[g_{m}nn^{H}g_{m}^{H}]=g_{m}\varepsilon[nn^{H}]g_{m}^{H}=g_{m}\sigma^{2}Ig_{m}^{H}

@@

as to $(gh)_{m}$, we have:

@@

\varepsilon[(gh)_{m}s_{d}s_{d}^{H}(gh)_{m}^{H}]=(gh)_{m}\varepsilon[s_{d}s_{d}^{H}](gh)_{m}^{H}=(gh)_{m}E_{s}I(gh)_{m}^{H} =E_{s}||(gh)_{m}||^{2}\label{equation03}

@@

In (\ref{equation03}), $E_{s}||(gh)_{m}||^{2}=E_{s}|g_{m}h_{m}|^{2}+E_{s}\sum_{j\neq m}|g_{m}h_{j}|^{2}$, $E_{s}|g_{m}h_{m}|^{2}$ is the post processing energy of detected symbol, $E_{s}\sum_{j\neq m}|g_{m}h_{j}|^{2}$ is the energy of spatial inter symbol interference. \\
Considering zero forcing algorithm first, the spatial equalizer $(\mathbf{G}_{j})_{ZF}=(\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}\mathbf{H}_{j}^{H}$, therefore from (\ref{equation0}), $\mathbf{G}_{j}\mathbf{H}_{j}=\mathbf{I}_{N_t-N}$ and

@@

Let $\omega_{m}$ denote the $m$th row of $(\mathbf{H_{j}^{H}H_{j}})^{-1}$, we have

@@

\parallel g_{m} \parallel^{2}=g_{m}g_{m}^{H}=\omega_{m}\mathbf{H}_{j}^{H}\mathbf{H}_{j}\omega_{m}^{H}\label{equation04}

@@

In (\ref{equation04}) we have $(\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}\mathbf{H}_{j}^{H}\mathbf{H}_{j}=\mathbf{I}_{N_t-N}$, hence $\omega_{m}\mathbf{H_{j}}^{H}\mathbf{H_{j}}=e_{m}$, where $e_{m}$ deontes the row vector with $m$th element equals to 1, the others are all 0. Therefore (\ref{equation04}) can be written as:

@@

\parallel g_{m} \parallel^{2}=((\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}_{mm})^{H}

@@

$((\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}_{mm})^{H}$ denotes the $m$th diagonal element of $((\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1})^{H}$. Because $\mathbf{H}_{j}^{H}\mathbf{H}_{j}$ is a Hermitian matrix, thus $((\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1})^{H}=(\mathbf{H}_{j}^{H}\mathbf{H}_{j})^{-1}$\cite{7} and (\ref{equation05}) can be written as:

@@

(\rho_{ZF}^{j})_{m}=\frac{E_{s}}{\sigma^{2}(\mathbf{H}_{j}^{H}\mathbf{H}_{j})_{mm}^{-1}}\label{equation02}

@@

Consider minimum mean square error (MMSE) detection, where $\mathbf{G}=\mathbf{G}_{MMSE}=(\mathbf{H}_{j}^{H}\mathbf{H}_{j}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}\mathbf{H}_{j}^{H}$, from \cite{5} \cite{6}, we have the expression of the post processing SINR of MMSE:

@@

(\rho_{MMSE}^{j})_{m}=\frac{E_{s}}{\sigma^{2}(\mathbf{H}_{j}^{H}\mathbf{H}_{j}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}_{mm}}-1\label{equation3}

@@

where $(\mathbf{H}_{j}^{H}\mathbf{H}_{j}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}_{mm}$ denotes the $m$th diagonal element of $(\mathbf{H}_{j}^{H}\mathbf{H}_{j}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$.
The criterion of the diversity maximization selection scheme is to choose the antenna subset which can maximize the worst post-processing SINR of the sub-datastream. The column elements of $\mathbf{H}_{j}$ are mutually independent Gaussian random variables, under favorable propagation condition, when $N_{r}\to \infty$ and the 2 norm of the column will become much larger than the inner product of two different columns. Hence, the columns of $\mathbf{H}_{j}$ become asymptotically orthogonal, a phenomenon that can be expressed as

@@

\lim_{N_{r}\to \infty}\frac{(\mathbf{H}_{j}^{H}\mathbf{H}_{j})}{N_{r}}=\mathbf{I}_{N_t-N}

@@

 Where $\mathbf{I}_{N_t-N}$ denotes an $(N_{t}-N)\times (N_{t}-N)$ identity matrix. Based on this property, for any $m$, $(\rho_{MMSE}^{j})_{m}$ and $(\rho_{ZF}^{j})_{m}$ in (\ref{equation02})(\ref{equation3}) are all the same, meaning that we can select arbitrarily the $N_{t}-N$ antenna subset.

@@

y=\mathbf{H}_{1}s_{1}+\mathbf{H}_{2}s_{2}+n\label{equation 4}

@@

Hence there are $N_{u}=$ $N_{t}\choose N$ possible antenna subsets, $\mathbf{H}_{1}\in \mathbb{C}^{N_{r}\times N}$ and $\mathbf{H_{2}}\in C^{N_{r}\times (N_{t}-N)}$.\\

@@

y^{k}=y-\mathbf{H}_{1}s_{1}^{k}   \label{equation41}

@@

$s_{1}^{k}$ denotes the $k$th possible $N\times 1$ symbol vector that is chosen in channel partition stage.
For ML detection:

@@

arg\min_{s_{1},s_{2}}\parallel y-\mathbf{H}_{1}s_{1}-\mathbf{H}_{2}s_{2} \parallel

@@

\textbf{3}. For every $s_{1}^{k}$ calculate $y_{k}=y-\mathbf{H}_{1}s_{1}^{k}$. There are $M^{N}$ possible $s_{1}^{k}$ as well as $y^{k}$. Using linear detection or OSIC to estimate $s_{2}^{k}$, the estimation is expressed as $\hat{s_{2}}^{k}$, then mapping $\hat{s_{2}}^{k}$ to the constellation to the nearest points, we use $s_{2}^{k}$ to represent it, every $s_{1}^{k}$ corresponds to one $s_{2}^{k}$, hence there are $M^{N}$ posssible $s_{2}^{k}$.\\

@@

\hat{s}=arg\min_{k=1,2,\ldots,M^{N}}\parallel y^{k}-\mathbf{H}_{2}s_{2}^{k} \parallel^{2}\label{equation6}

@@

According to section \ref{channel partition}, based on diversity maximization antenna selection criterion and channel hardening phenomenon we can choose an arbitrary antenna subset, and hence the cost of antenna selection is zero. 
The second step calculating $y^{k}=y-\mathbf{H}_{1}s_{1}^{k}$, including an matrix multiplication and vector substruction,

@@

 complex substitution are required. Then calculate the MMSE estimation $\hat{s}_{2}^{k}$:

@@

\hat{s}_{2}^{k}=[\mathbf{H}_{2}^{H}\mathbf{H}_{2}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H}_{2}^{H}y^{k}\label{equation7}

@@

$\mathbf{H}_{2}\in \mathbb{C}^{N_{r}\times (N_{t}-N)}$, the calculation of weight matrix $\mathbf{G}_{MMSE}=[\mathbf{H}_{2}^{H}\mathbf{H}_{2}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H}_{2}^{H}$, because of the channel hardening phenomenon mentioned in section \ref{channel partition}, $\mathbf{G}_{MMSE}=[\mathbf{H}_{2}^{H}\mathbf{H}_{2}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H}_{2}^{H}=\frac{1}{N_{r}+\frac{\sigma^{2}}{E_{s}}}\mathbf{I}\mathbf{H}_{2}^{H}$, hence the calculation of $\mathbf{G}_{MMSE}$ is real-complex multiplication, which cost 2 flops, 1/3 of complex multiplication, hence get $\mathbf{G}_{MMSE}$ requires:

@@

complex multiplication.\\
In the whole process $\mathbf{G}_{MMSE}$ only need to be calculated one time. Calculation of $\hat{s}_{2}^{k}=\mathbf{G}_{MMSE}y_{k}$ is the complex matrix multiplication which requires:

@@

M^{N}(N_rN+(N_{t}-N)N_r+N_r(N_t-N)+\frac{N_r}{2})+\frac{(N_{t}-N)N_{r}}{3}=M^{N}(2N_{r}N_{t}-N_{r}N+\frac{N_r}{2})+\frac{(N_{t}-N)N_r}{3}\label{equation8}

@@

f_{sel-MMSE}(N) &=& 6[M^{N}(2N_{r}N_{t}-N_{r}N+\frac{N_r}{2})+\frac{(N_{t}-N)N_r}{3}]\\  

@@

where $N_{u}={\displaystyle{N_{t}\choose N}}$. Table \ref{table_sel-MMSE} shows the complexity comparison between $f_{sel-MMSE}$and$f_{sel-MMSE_{O}}$, at channel partition stage they all use diversity maximization scheme in \cite{2} to determine $N$.
\begin{table}[t]
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\diagbox{$M$}{$\frac{f_{sel-MMSE}}{f_{sel-MMSE_{O}}}$}{ $N_{r}\times N_{t}$}&$10\times 10$&$32\times 32$&$64\times 64$&$128\times 128$\\\hline

BPSK&0.0249&1.3746e-05&2.3406e-09&1.5672e-14\\\hline

QPSK&0.1683&4.3817e-04&5.9893e-07&3.2095e-11\\\hline

16QAM&0.9282&0.3098&0.0378&1.3460e-04\\\hline

64QAM&0.9988&0.9978&0.9996&0.9982\\
\hline
\end{tabular}
\caption{complexity comparison between $f_{sel-MMSE}$ and $f_{sel-MMSE_{O}}$}
\label{table_sel-MMSE}
\end{table} 
From table \ref{table_sel-MMSE} we can see, as the size of array increases, the complexity abbreviation gain increases, as the size of signal constellation increases, the complexity abbreviation gain decreases.

@@

First we use a general case to illustrate how MMSE-OSIC works. Consider a MIMO system as mentioned in section \ref{system}:

@@

With $\mathbf{G}_{MMSE}=(\mathbf{H}^{H}\mathbf{H}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}\mathbf{H}^{H}$, after equalization using $\mathbf{G}_{MMSE}$,  (\ref{equation07}) can be changed to 

@@

\hat{s}=\mathbf{G}_{MMSE}y=\mathbf{G}_{MMSE}Hs+\mathbf{G}_{MMSE}n\label{equation08}

@@

where $\hat{s}$ denote the unconstrained estimate of $s$. From (\ref{equation08}) we can express $m$th estimation of sub-datastream $\hat{s}_{m}$:

@@

\hat{s}_{m}={g_{m}}y\label{equation09}

@@

where $g_{m}$ denote the $m$th row of $\mathbf{G}_{MMSE}$, unlike MMSE, the MMSE-OSIC algorithm detects the $s_{m}$ symbol by symbol. The detection order at every iteration is based on the following principle: $\hat{s_{m}}$ is detected based on their post-processing SINR at each iteration, where $\hat{s}_{m}$ with the highest SNR will be detected. According to (\ref{equation3}), the post processing SINR is associated to the diagonal elements of $(\mathbf{H}^{H}\mathbf{H}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$. The smaller is the $m$th diagonal element of $(\mathbf{H}^{H}\mathbf{H}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$, the higher is the post-processing SINR of $\hat{s}_{m}$. Because there are $N_{t}$ symbols in vector $s$, hence MMSE-OSIC algorithm will detect $s_{m}$ for $N_{t}$ times, in the $i$th ($i\in [0,N_{t}-1]$) iteration, it majorly has 3 steps:\\
\textbf{1}.Ordering the channel based on the principle mentioned above, choose the sub-datastream with highest SNR(smallest diagonal element of $((\mathbf{H}_{j}^{i})^{H}\mathbf{H}_{j}^{i}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$ to detect, $\mathbf{H}_{j}^{i}\in \mathbb{C}^{Nr\times (Nt-N-i)}$ denotes the $i$th step sub-matrix in MMSE-OSIC iteration from the $j$th propagation matrix after channel partition as mentioned in section \ref{channel partition}, $\mathbf{H}_{j}^{0}=\mathbf{H}_{j}$\\ 
\textbf{2}.Find the unconstrained estimate

@@

\hat{s}_{i+1}=g_{i+1}y_{i}\label{equation012}

@@

and slice $\widehat{s_{i+1}}$ to $\bar{s}_{i+1}$ corresponds to signal constellation.\\
\textbf{3}.Cancel the information correlated to  $\bar{s_{i}}$ from received vector $y_{i}$, $h_{m}\quad m\in [1,N_{t}]$ denotes the $m$th column of $\mathbf{H}_{j}^{i}$, $s_{2m}$ denotes the $m$th symbol of $s_{2}$. We have

@@

y_{k}=\sum_{m=1,2,\ldots Nt-N}h_{m}s_{2m}+n

@@

Then update propagation matrix by removing the column $h_{i+1}$ from $\mathbf{H}_{j}^{i}$.\\

>>>>

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
\documentclass[12pt,a4paper,final]{article}
% Add the compsoc option for Computer Society conferences.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
\usepackage[latin1]{inputenc}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
\usepackage{psfrag}
\usepackage{color}
\usepackage{epsfig}
\usepackage{appendix}
\usepackage{caption}
\usepackage{cite}
\usepackage{ifpdf}
\usepackage[cmex10]{amsmath}
\usepackage{algorithmic}
\usepackage{array}
\usepackage{stfloats}
\usepackage{url}
\usepackage{fixltx2e}
\usepackage{setspace}

@@

%
% paper title
% can use linebreaks \\ within to get better formatting as desired
% Do not put math or special symbols in the title.
\title{Parallel Acceleration of Fixed Complexity Sphere Decoder for Large MIMO Uplink System Based on CUDA}

@@

\section{Introduction}

\section{System Model}\label{system}
% no \IEEEPARstart
%This demo file is intended to serve as a ``starter file''
%for IEEE conference papers produced under \LaTeX\ using
%IEEEtran.cls version 1.8 and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
%I wish you the best of success.
%\hfill mds
%\hfill December 27, 2012
Consider a MIMO system with $N_r$ receive antennas and $N_t$ transmit antennas, where $N_r$ is large, for simplification, we assume $N_r\to\infty$, the corresponding discrete time model is given:
\begin{equation}
y=\mathbf{H}s+n
\end{equation}
$s\in \mathbb{C}^{N_{t}\times 1}$ is the transmit symbol vector, $y\in \mathbb{C}^{N_{r}\times 1}$ is the receive symbol vector, $\mathbf{H}\in \mathbb{C}^{N_{r}\times N_{t}}$ is the channel propagation matrix with independent identity distributed(i.i.d) circularly symmetric complex Gaussian zero mean elements of unit variance (Rayleigh fading), $n\in \mathbb{C}^{N_{r}\times 1}$ is the additive white Gaussian noise (AWGN) with zero mean elements of variance $\sigma^{2}$, $E(ss^{H})=E_{s}\mathbf{I}_{Nt}$, $E_s$ denotes the average symbol energy, $E$ denotes the expectation operation and $M$ denotes the constellation size.

@@

y_{d}=\mathbf{H_{j}}s_{d}+n

@@

\hat{s_{d}}=\mathbf{G}_{j}y_{d}=\mathbf{G_{j}H_{j}}s_{d}+\mathbf{G}_{j}n\label{equation0}

@@

[\hat{s_{d}}]_{m}=(\mathbf{G}_{j}\mathbf{H_{j}})_{m}s_{d}+g_{m}n\label{equation01}

@@

where $[\hat{s_{d}}]_{m}$ denotes the $m$th element of $\hat{s_{d}}$, $(\mathbf{G}_{j}\mathbf{H_{j}})_{m}$ denotes the $m$th row of $\mathbf{GH_{j}}$, $g_{m}$ denotes the $m$th row of $\mathbf{G}_{j}$ and $h_{m}$ denotes the $m$th column of $\mathbf{H_{j}}$, The post processing SNR of $m$th sub datastream and $j$th selected antenna subset is defined as\cite{3}:

@@

E[|g_{m}n|^{2}]= E[(g_{m}n)(g_{m}n)^{H}]=E[g_{m}nn^{H}g_{m}^{H}]=g_{m}E[nn^{H}]g_{m}^{H}=g_{m}\sigma^{2}Ig_{m}^{H}

@@

as to $(\mathbf{G}_{j}\mathbf{H_{j}})_{m}$, we have:

@@

E((\mathbf{G}_{j}\mathbf{H_{j}})_{m}s_{d}s_{d}^{H}(\mathbf{G}_{j}\mathbf{H}_{j})_{m}^{H})=(\mathbf{G}_{j}\mathbf{H}_{j})_{m}E[s_{d}s_{d}^{H}](\mathbf{G}_{j}\mathbf{H}_{j})_{m}^{H}=(\mathbf{G}_{j}\mathbf{H}_{j})_{m}E_{s}I(\mathbf{G}_{j}\mathbf{H}_{j})_{m}^{H} =E_{s}||(\mathbf{G}_{j}\mathbf{H}_{j})_{m}||^{2}\label{equation03}

@@

In (\ref{equation03}),$E_{s}||(\mathbf{G}_{j}\mathbf{H}_{j})_{m}||^{2}=E_{s}|g_{m}h_{m}|^{2}+E_{s}\sum_{j\neq m}|g_{m}h_{j}|^{2}$, $E_{s}|g_{m}h_{m}|^{2}$ is the post processing energy of detected symbol, $E_{s}\sum_{j\neq m}|g_{m}h_{j}|^{2}$ is the energy of spatial inter symbol interference. \\
Considering zero forcing algorithm first, the spatial equalizer $\mathbf{G_{j}}_{ZF}=(\mathbf{H_{j}}^{H}\mathbf{H_{j}})^{-1}\mathbf{H_{j}}^{H}$, therefore from (\ref{equation0}) $\mathbf{G}_{j}\mathbf{H}_{j}=\mathbf{I}_{N_t-N}$ and

@@

Let $(\mathbf{H_{j}^{H}H_{j}})^{-1}_{m}$ denote the $m$th row of $(\mathbf{H_{j}^{H}H_{j}})^{-1}$, we have

@@

\parallel g_{m} \parallel^{2}=g_{m}g_{m}^{H}=(\mathbf{H_{j}^{H}H_{j}})^{-1}_{m}\mathbf{H_{j}}^{H}\mathbf{H_{j}}((\mathbf{H_{j}^{H}H_{j}})^{-1}
_{m})^{H}\label{equation04}

@@

In (\ref{equation04}) we use $(\mathbf{H_{j}^{H}H_{j}})^{-1}\mathbf{H_{j}}^{H}\mathbf{H_{j}}=\mathbf{I}_{N_t-N}$ hence $(\mathbf{H_{j}^{H}H_{j}})^{-1}_{m}\mathbf{H_{j}}^{H}\mathbf{H_{j}}=e_{m}$, where $e_{m}$ deontes the row vector with $m$th element equals to 1, the others are all 0. Therefore (\ref{equation04}) can be written as:

@@

\parallel g_{m} \parallel^{2}=((\mathbf{H_{j}}^{H}\mathbf{H_{j}})^{-1}_{mm})^{H}

@@

$((\mathbf{H_{j}}^{H}\mathbf{H_{j}})^{-1}_{mm})^{H}$ denotes the $m$th diagonal element of $((\mathbf{H_{j}}^{H}\mathbf{H_{j}})^{-1})^{H}$. Because $\mathbf{H_{j}^{H}}\mathbf{H_{j}}$ is a Hermitian matrix, thus $((\mathbf{H_{j}}^{H}\mathbf{H_{j}})^{-1})^{H}=(\mathbf{H_{j}}^{H}\mathbf{H_{j}})^{-1}$\cite{7} and (\ref{equation05}) can be written as:

@@

(\rho_{ZF}^{j})_{m}=\frac{E_{s}}{\sigma^{2}(\mathbf{H_{j}}^{H}\mathbf{H_{j}})_{mm}^{-1}}\label{equation02}

@@

Consider minimum mean square error (MMSE) detection, where $\mathbf{G}=\mathbf{G_{MMSE}}=(\mathbf{H_{j}}^{H}\mathbf{H_{j}}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}\mathbf{H_{j}}^{H}$, from \cite{5} \cite{6}, we have the expression of the post processing SNR of MMSE:

@@

(\rho_{MMSE}^{j})_{m}=\frac{E_{s}}{\sigma^{2}(\mathbf{H_{j}}^{H}\mathbf{H_{j}}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}_{mm}}-1\label{equation3}

@@

where $(\mathbf{H_{j}}^{H}\mathbf{H_{j}}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}_{mm}$ denotes the $m$th diagonal element of $(\mathbf{H_{j}}^{H}\mathbf{H_{j}}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$.
The criterion of the diversity maximization selection scheme is to choose the antenna subset which can maximize the worst post-processing SNR of the sub-datastream. The column elements of $\mathbf{H_{j}}$ are mutually independent Gaussian random variables under favorable propagation condition. When $N_{r}\to \infty$ and the 2 norm of the column will become much larger than the inner product of two different columns, hence, the columns of $\mathbf{H_{j}}$ become asymptotically orthogonal, a phenomenon that can be expressed as

@@

\lim_{N_{r}\to \infty}\frac{(\mathbf{H_{j}}^{H}\mathbf{H_{j})}}{N_{r}}=\mathbf{I}_{N_t-N}

@@

 Where $\mathbf{I}_{N_t-N}$ denotes an $(N_{t}-N)\times (N_{t}-N)$ identity matrix. Based on this property, for any $m$, $(\rho_{MMSE}^{j})_{m}$ and $(\rho_{ZF}^{j})_{m}$ in (\ref{equation02})(\ref{equation3}) are all the same, means that we can select arbitrarily the $N_{t}-N$ antenna subset.

@@

y=\mathbf{H_{1}}s_{1}+\mathbf{H_{2}}s_{2}+n\label{equation 4}

@@

Hence there are $N_{u}=$ $N_{t}\choose N$ possible antenna subsets,$\mathbf{H_{1}}\in \mathbb{C}^{N_{r}\times N}$ and $\mathbf{H_{2}}\in C^{N_{r}\times (N_{t}-N)}$,\\

@@

y^{k}=y-\mathbf{H_{1}}s_{1}^{k}   \label{equation41}

@@

$s_{1}^{k}$ denotes the $k$th possible $N\times 1$ symbol vector that chosen in channel partition stage.
for ML detection:

@@

arg\min_{s}\parallel y-\mathbf{H_{1}}s_{1}-\mathbf{H_{2}}s_{2} \parallel

@@

\textbf{3}. For every $s_{1}^{k}$ calculate $y_{k}=y-\mathbf{H_{1}}s_{1}^{k}$, There are $M^{N}$ possible $s_{1}^{k}$ as well as $y_{k}$, Using linear detection or OSIC to estimate $s_{2}^{k}$, the estimation is expressed as $\hat{s_{2}}^{k}$, then mapping $\hat{s_{2}}^{k}$ to the constellation to the nearest points, we use $s_{2}^{k}$ to represent it, every $s_{1}^{k}$ corresponds to one $s_{2}^{k}$, hence there are $M^{N}$ posssible $s_{2}^{k}$.\\

@@

\hat{s}=arg\min_{k=1,2,\ldots,N_{u}}\parallel y_{k}-\mathbf{H_{2}}s_{2}^{k} \parallel^{2}\label{equation6}

@@

According to section \ref{channel partition}, based on diversity maximization antenna selection criterion we can choose an arbitrary antenna subset, and hence the cost of antenna selection is zero. 
The second step calculating $y_{k}=y-\mathbf{H_{1}}s_{1}^{k}$, including an matrix multiplication and vector substruction,

@@

 complex additions and substitution are required. Then calculate the MMSE estimation $\hat{s}_{2}^{k}$:

@@

\hat{s}_{2}^{k}=[\mathbf{H_{2}}^{H}\mathbf{H_{2}}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H_{2}}^{H}y_{k}\label{equation7}

@@

$\mathbf{H_{2}}\in \mathbb{C}^{N_{r}\times (N_{t}-N)}$, the calculation of weight matrix $\mathbf{G_{MMSE}}=[\mathbf{H_{2}}^{H}\mathbf{H_{2}}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H_{2}}^{H}$, because of the channel hardening phenomenon mentioned in section \ref{channel partition}, $\mathbf{G_{MMSE}}=[\mathbf{H_{2}}^{H}\mathbf{H_{2}}+\frac{\sigma^{2}}{E_{s}}\mathbf{I}]^{-1}\mathbf{H_{2}}^{H}=\frac{1}{N_{r}+\frac{\sigma^{2}}{E_{s}}}\mathbf{I}\mathbf{H_{2}}^{H}$, hence the calculation of $\mathbf{G_{MMSE}}$ is real-complex multiplication, which cost 2 flops, 1/3 of complex multiplication, hence get $\mathbf{G_{MMSE}}$ requires:

@@

In the whole process (\ref{equation31}) only need to be calculated one time. Calculation of $\hat{s}_{2}^{k}=\mathbf{G_{MMSE}}y_{k}$ is the complex matrix multiplication which requires:

@@

M^{N}(N_rN+(Nt_-N)N_r+N_r(N_t-N)+\frac{N_r}{2})+\frac{(N_{t}-N)N_{r}}{3}=M^{N}(2N_{r}N_{t}-N_{r}N+\frac{N_r}{2})+\frac{(N_{t}-N)N_r}{3}\label{equation8}

@@

f_{sel-MMSE}(N) &=& 6(M^{N}(2N_{r}N_{t}-N_{r}N+\frac{N_r}{2})+\frac{(N_{t}-N)N_r}{3})\\  

@@

@@

First we use a general case to illustrate how MMSE-OSIC works, first consider a MIMO system as mentioned in section \ref{system}:

@@

With $\mathbf{G}_{MMSE}=(\mathbf{H}^{H}\mathbf{H}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}\mathbf{H}^{H}$, after equalization using $\mathbf{G}_{MMSE}$ (\ref{equation07}) can be changed to 

@@

\hat{s}=\mathbf{G_{MMSE}}y=\mathbf{G_{MMSE}H}s+\mathbf{G_{MMSE}}n\label{equation08}

@@

where $\hat{s}$ denotes the unconstrained estimation of $s$. From (\ref{equation08}) we can express $m$th estimation of sub-datastream $\hat{s}_{m}$:

@@

\hat{s}_{m}={g^{T}_{m}}y\label{equation09}

@@

where $g_{m}$ denotes the $m$th row of $\mathbf{G_{MMSE}}$, unlike MMSE, MMSE-OSIC algorithm detects the $s_{m}$ symbol by symbol. The detection order at every iteration is based on the following principle: $\hat{s_{m}}$ is detected based on their post-processing SNR at each iteration, $\hat{s}_{m}$ with the highest SNR will be detected , according to (\ref{equation3}), this element is associated to the diagonal elements of $(\mathbf{H}^{H}\mathbf{H}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$. The smaller is the $m$th diagonal element of $(\mathbf{H}^{H}\mathbf{H}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$, the higher is the post-processing SNR of $\hat{s}_{m}$. because there are $N_{t}$ symbols in vector $s_{m}$, hence MMSE-OSIC algorithm will detect $s_{m}$ for $N_{t}$ times, in the $i$th ($i\in [0,N_{t}-1]$) iteration, it majorly has 3 steps:\\
\textbf{1}.Ordering the channel based on the principle mentioned above, choose the sub-datastream with highest SNR(smallest diagonal element of $(\mathbf{H_{j}^{i}}^{H}\mathbf{H_{j}^{i}}+\frac{\sigma^{2}}{E_{s}}\mathbf{I})^{-1}$ to detect, $H_{j}^{i}\in \mathbb{C}^{Nr\times Nt-N-i}$ denotes the $i$th step sub-matrix in MMSE-OSIC iteration from the $j$th propagation matrix after channel partition as mentioned in section \ref{channel partition}, $\mathbf{H_{j}^{0}}=\mathbf{H_{j}}$\\ 
\textbf{2}.Find the unconstrained estimation

@@

\hat{s}_{i+1}=g_{i+1}^{T}y_{i}\label{equation012}

@@

and slice $\widehat{s_{i+1}}$ to $\bar{s}_{i+1}$ corresponding to symbol constellation.\\
\textbf{3}.Cancel the information correlated to  $\bar{s_{i}}$ from received vector $y_{i}$, $h_{m} m\in [1,N_{t}-N]$ denotes the $m$th column of $\mathbf{H_{j}^{i}}$. Hence from (\ref{equation07}), we have

@@

y_{k}=\sum_{m=1,2,\ldots Nt-N}h_{m}s_{2}^{m}+n

@@

Then update propagation matrix by removing the column $h_{i+1}$ from $\mathbf{H_{j}^{i}}$.\\
